CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-07-07 script/de/chapters/numbertheory.tex
*** empty log message ***

Mit $J(5) = 4$ ergibt sich:
$$
(4^3)^2 \equiv 4^{3*2{\rm ~(mod~}J(5))} \equiv 4^{6{\rm ~mod~}4} \equiv 4^2 \equiv 16 \equiv 1 {\rm ~(mod~}5).
$$
Das liefert das richtige Ergebnis.
\end{quote}
\vskip + 10 pt

@@

{\bf Beispiel 2:}
$$
3^{28} \equiv 3^{4*7} \equiv 3^{4*7{\rm ~(mod~}10)} \equiv 3^8 \equiv 6561 \equiv 5 {\rm ~(mod~}11).
$$

@@

Die Berechnungen dazu fanden im Bundesamt f"ur Sicherheit in der 
Informationstechnik (BSI) in Bonn statt\footnote{%
Das BSI \index{BSI} erstellt jedes Jahr ein Papier "uber die Eignung von
Kryptoalgorithmen, mit denen Signaturen erzeugt werden k"onnen,         
die den Vorgaben des deutschen Signaturgesetzes gen"ugen.
Bei dieser Erstellung werden Experten aus Wirtschaft und Wissenschaft
beteiligt. Um die Eignung von Signaturverfahren zu beurteilen, deren
Schwierigkeit auf dem Faktorisierungsproblem beruht,
kooperiert das BSI auch mit Forschern der Universit"at Bonn.
Weitere Informationen zu Kryptoalgorithmen auf den BSI-Internetseiten unter:
   \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
{\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}}
}.

>>>>

Mit $J(5) = 4$ ergibt sich:
$$
(4^3)^2 \equiv 4^{3\:*\:2{\rm ~(mod~}J(5))} \equiv 4^{6{\rm ~mod~}4} \equiv 4^2 \equiv 16 \equiv 1 {\rm ~(mod~}5).
$$
Das liefert das richtige Ergebnis.
\end{quote}
\vskip + 10 pt

@@

{\bf Beispiel 2:}
$$
3^{28} \equiv 3^{4{*}7} \equiv 3^{4{*}7{\rm ~(mod~}10)} \equiv 3^8 \equiv 6561 \equiv 5 {\rm ~(mod~}11).
$$

@@

Die Berechnungen dazu fanden im Bundesamt f"ur Sicherheit in der 
Informationstechnik (BSI) in Bonn statt\footnote{%
Das BSI \index{BSI} erstellt jedes Jahr ein Papier "uber die Eignung von
Kryptoalgorithmen, mit denen Signaturen erzeugt werden k"onnen,         
die den Vorgaben des deutschen Signaturgesetzes gen"ugen.
Bei dieser Erstellung werden Experten aus Wirtschaft und Wissenschaft
beteiligt. Um die Eignung von Signaturverfahren zu beurteilen, deren
Schwierigkeit auf dem Faktorisierungsproblem beruht,
kooperiert das BSI auch mit Forschern der Universit"at Bonn.
Weitere Informationen zu Kryptoalgorithmen finden Sie auf den BSI-Internetseiten unter:
   \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
{\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}}
}.

### 2003-07-07 script/de/chapters/primes.tex
*** empty log message ***

Nahezu alle bekannten riesig gro"sen Primzahlen sind spezielle 
Kandidaten, sogenannte \index{Mersenne Marin}{\em Mersennezahlen}\footnote{%
Marin Mersenne, franz"osischer Priester und Mathematiker,
08.09.1588 -- 01.09.1648.
\index{Mersenne Marin}
}
der Form $2^p -1,$ wobei $p$ eine Primzahl ist.
Nicht alle Mersennezahlen sind prim:
$$
\begin{array}{cl}
2^2 - 1 = 3 & \Rightarrow {\rm prim} \\
2^3 - 1 = 7 & \Rightarrow {\rm prim} \\
2^5 - 1 = 31    & \Rightarrow {\rm prim} \\
2^7 - 1 = 127    & \Rightarrow {\rm prim} \\
2^{11} - 1 = 2.047 = 23 \cdot 89    & \Rightarrow  {\rm NICHT~prim} !
\end{array}
$$

@@

Mersenne behauptete, dass $2^{67}-1$ eine Primzahl ist. Auch zu
dieser Behauptung gibt es eine interessante mathematische Historie:
Zuerst dauerte es "uber 200 Jahre, bis \index{Lucas Edouard}
Edouard Lucas (1842-1891) bewies, dass diese Zahl zusammengesetzt
ist. Er argumentierte aber indirekt und kannte keinen der
Faktoren. Dann zeigte Frank Nelson Cole 1903, aus welchen Faktoren
diese Primzahl besteht:
$$ 2^{67} -1 =147. 573. 952. 589. 676. 412. 927 = 193. 707. 721 \cdot 761. 838. 257. 287. $$
Er gestand, 20 Jahre an der Faktorisierung 
\index{Faktorisierung} (Zerlegung in Faktoren)%
\footnote{%
Mit CrypTool\index{CrypTool} v1.3 k"onnen Sie Zahlen auf folgende Weise 
faktorisieren: Men"u {\bf Einzelverfahren \textbackslash{} RSA-Demo 
\textbackslash{} Faktorisieren einer Zahl}.
}
dieser 21-stelligen Dezimalzahl gearbeitet zu haben!

@@

Schon im 19. Jahrhundert wusste man, dass die $29$-stellige Zahl
$$ f(7) = 2^{2^7} + 1 $$
keine Primzahl ist. Aber erst 1970 fanden Morrison/Billhart ihre Zerlegung.
\begin{eqnarray*}
f(7) & = & 340.282.366.920.938.463.463.374.607.431.768.211.457 \\
& = & 59. 649. 589. 127. 497. 217 \cdot  5.704.689.200.685.129.054.721
\end{eqnarray*}\label{f7}

@@

% --------------------------------------------------------------------------
\subsubsection[Vermutung von Catalan]{Vermutung von Catalan\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, belgischer Mathematiker, 30.5.1814$-$14.2.1894.\\
Nach ihm sind auch die sogenannten {\em Catalanzahlen}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2$ \\
$=  1, 2, 5, 14, 42, 132, 429, 1.430, 
4.862, 16.796, 58.786, 208.012, 742.900, 2.674.440, 9.694.845, ... $
benannt.
    }
Catalan\index{Catalan Eugene}\index{Zahlen!Catalanzahl}
"au"serte die Vermutung, dass $ C_4 $ eine Primzahl ist:
$$
\begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
\begin{sloppypar}
    (siehe
        {\href{http://www.utm.edu/research/primes/mersenne.shtml}{\tt http://www.utm.edu/research/primes/mersenne.shtml}}
        unter Conjectures and Unsolved Problems).
\end{sloppypar}

>>>>

Nahezu alle bekannten riesig gro"sen Primzahlen sind spezielle 
Kandidaten, sogenannte \index{Mersenne Marin}{\em Mersennezahlen}\footnote{%
Marin Mersenne, franz"osischer Priester und Mathematiker,
08.09.1588$-$01.09.1648.
\index{Mersenne Marin}
}
der Form $2^p -1,$ wobei $p$ eine Primzahl ist.
Nicht alle Mersennezahlen sind prim:
$$
\begin{array}{cl}
2^2 - 1 = 3 & \Rightarrow {\rm prim} \\
2^3 - 1 = 7 & \Rightarrow {\rm prim} \\
2^5 - 1 = 31    & \Rightarrow {\rm prim} \\
2^7 - 1 = 127    & \Rightarrow {\rm prim} \\
2^{11} - 1 = 2.047 = 23 \cdot 89    & \Rightarrow  {\rm NICHT~prim} !
\end{array}
$$

@@

Mersenne behauptete, dass $2^{67}-1$ eine Primzahl ist. Auch zu
dieser Behauptung gibt es eine interessante mathematische Historie:
Zuerst dauerte es "uber 200 Jahre, bis \index{Lucas Edouard}
Edouard Lucas (1842-1891) bewies, dass diese Zahl zusammengesetzt
ist. Er argumentierte aber indirekt und kannte keinen der
Faktoren. Dann zeigte Frank Nelson Cole\index{Cole Frank Nelson}\footnote{%
Frank Nelson Cole, amerikanischer Mathematiker, 20.09.1861$-$26.05.1926.} 1903,
aus welchen Faktoren diese Primzahl besteht:
$$ 2^{67} -1 =147. 573. 952. 589. 676. 412. 927 = 193. 707. 721 \cdot 761. 838. 257. 287. $$
Er gestand, 20 Jahre an der Faktorisierung 
\index{Faktorisierung} (Zerlegung in Faktoren)%
\footnote{%
Mit CrypTool\index{CrypTool} v1.3 k"onnen Sie Zahlen auf folgende Weise 
faktorisieren: Men"u {\bf Einzelverfahren \textbackslash{} RSA-Demo 
\textbackslash{} Faktorisieren einer Zahl}.
}
dieser 21-stelligen Dezimalzahl gearbeitet zu haben!

@@

Schon im 19. Jahrhundert wusste man, dass die $29$-stellige Zahl
$$ f(7) = 2^{2^7} + 1 $$
keine Primzahl ist. Aber erst 1970 fanden Morrison/Billhart ihre Zerlegung.
\begin{eqnarray*}\label{f7}
f(7) & = & 340.282.366.920.938.463.463.374.607.431.768.211.457 \\
& = & 59. 649. 589. 127. 497. 217 \cdot  5.704.689.200.685.129.054.721
\end{eqnarray*}

@@

% --------------------------------------------------------------------------
\subsubsection[Vermutung von Catalan]{Vermutung von Catalan\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, belgischer Mathematiker, 30.5.1814$-$14.2.1894.\\
Nach ihm sind auch die sogenannten {\em Catalanzahlen}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2$ \\
$=  1, 2, 5, 14, 42, 132, 429, 1.430, 
4.862, 16.796, 58.786, 208.012, 742.900, 2.674.440, 9.694.845, ... $
benannt.
    }
Catalan\index{Catalan Eugene}\index{Zahlen!Catalanzahl}
"au"serte die Vermutung, dass $ C_4 \;$ eine Primzahl ist:
$$
\begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
\begin{sloppypar}
    (siehe
        {\href{http://www.utm.edu/research/primes/mersenne.shtml}{\tt http://www.utm.edu/research/primes/mersenne.shtml}}
        unter Conjectures and Unsolved Problems).
\end{sloppypar}

### 2003-07-07 script/en/chapters/cryptomethods.tex
*** empty log message ***

The mathematicians Courtois and Pieprzyk \cite{Courtois2002} published a paper
in 2002, which got a great deal of attention amongst the crypto community: The
pair had further developed the XL-method (eXtended Linearization), introduced at
Eurocrypt 2000 by Shamir et al., to create the so called XSL-method (eXtended
Sparse Linearization). The XL-method is a heuristic technique, which in some
cases manages to solve big non-linear equation systems and which was till then
used to analyze an asymmetric algorithm (HFE).  The innovation of Courtois and
Pieprzyk was, to apply the XL-method on symmetric codes: the XSL-method can be
applied to very specific equation systems. A 256-bit AES could be attacked in
roughly $2^{230}$ steps. This is still a purely academic attack, but also a
direction pointer for a complete class of block ciphers. The major problem with
this attack is that until now nobody has worked out, under what conditions it is
successful: the authors specify in their paper necessary conditions, but it is
not known, which conditions are sufficient.  There are two very new aspects of
this attack: firstly this attack is not based on statistics but on algebra. So
attacks seem to be possible, where only very small amounts of cipher text are
available. Secondly the security of a product-algorithm\index{Cascade cipher}%
\footnote{%
A cipher text can be used as input for another encryption algorithm. 
A cascade cipher is build up as a composition of different encryption 
transformations.\\
Cascading does not always improve the security.\\
This process is also used {\em within} modern algorithms:
They usually combine simple and, considered at its own, cryptologically
relatively unsecure single steps in several rounds into an efficient 
overall procedure.  Most block ciphers (e.g. DES, IDEA) are cascade ciphers.\\
Also serial usage of the same cipher with different keys (like with Triple-DES)
is called cascade cipher. 
}
does not exponentially increase with the number of rounds.

>>>>

The mathematicians Courtois and Pieprzyk \cite{Courtois2002} published a paper
in 2002, which got a great deal of attention amongst the crypto community: The
pair had further developed the XL-method (eXtended Linearization), introduced at
Eurocrypt 2000 by Shamir et al., to create the so called XSL-method (eXtended
Sparse Linearization). The XL-method is a heuristic technique, which in some
cases manages to solve big non-linear equation systems and which was till then
used to analyze an asymmetric algorithm (HFE).  The innovation of Courtois and
Pieprzyk was, to apply the XL-method on symmetric codes: the XSL-method can be
applied to very specific equation systems. A 256-bit AES could be attacked in
roughly $2^{230}$ steps. This is still a purely academic attack, but also a
direction pointer for a complete class of block ciphers. The major problem with
this attack is that until now nobody has worked out, under what conditions it is
successful: the authors specify in their paper necessary conditions, but it is
not known, which conditions are sufficient.  There are two very new aspects of
this attack: firstly this attack is not based on statistics but on algebra. So
attacks seem to be possible, where only very small amounts of cipher text are
available. Secondly the security of a product algorithm\index{Cascade cipher}%
\footnote{%
A cipher text can be used as input for another encryption algorithm. 
A cascade cipher is build up as a composition of different encryption 
transformations. The overall cipher is called product algorithm or cacade cipher (somtimes depending whether the used keys are statistically dependent or not).\\
Cascading does not always improve the security.\\
This process is also used {\em within} modern algorithms:
They usually combine simple and, considered at its own, cryptologically
relatively unsecure single steps in several rounds into an efficient 
overall procedure.  Most block ciphers (e.g. DES, IDEA) are cascade ciphers.\\
Also serial usage of the same cipher with different keys (like with Triple-DES)
is called cascade cipher. 
}
does not exponentially increase with the number of rounds.

### 2003-07-07 script/en/chapters/digitalsignatures.tex
*** empty log message ***

\paragraph{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random\index{Random}, less than $q$\\
$r$ & $= (g^k \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source\index{Random}  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function.
\end{itemize}
\paragraph{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1}  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\

@@

\bibitem[X.509v3]{X.509v3} ITU-T, \\
    \index{X.509} \index{ITU-T} \index{ISO/IEC 9594-8}
    {\em X.509 (1993) Amendment 1: Certificate Extensions, The Directory
    Authentication Framework},\\ 
    International Telecommunication Union, Geneva, Switzerland, July 1995\\
    (equivalent to Amendment 1 to ISO/IEC 9594-8).

@@

>>>>

\paragraph{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random\index{Random}, less than $q$\\
$r$ & $= (g^k \; \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source\index{Random}  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function.
\end{itemize}
\paragraph{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1} \;  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\

@@

@@

\bibitem[X.509v3]{X.509v3} ITU-T, \\
    \index{X.509} \index{ITU-T} \index{ISO/IEC 9594-8}
    {\em X.509 (1993) Amendment 1: Certificate Extensions, The Directory
    Authentication Framework},\\ 
    International Telecommunication Union, Geneva, Switzerland, July 1995\\
    (equivalent to amendment 1 to ISO/IEC 9594-8).

### 2003-07-07 script/en/chapters/numbertheory.tex
*** empty log message ***

The  computations for the factorization of RSA-160 took place at the 
German Information Security Agency (GISA) in Bonn\footnote{%
Every year the GISA\index{GISA} creates a paper to describe which 
crypto algorithms are feasible to generate digital signatures according
to the German signature law -- under participation of experts from
economy and science. To review signature methods based on the
factorisation problem the GISA also co-operates with researchers from the
University of Bonn.
Further information about crypto algorithms on the web page of GISA:
   \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
{\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}}.
}.

@@

  \item {Mr.\ Henrik Koy for making many very useful suggestions, 
         for the very constructive proof-reading this article and
         for helping with TeX. \\
	 Mr.\ Koy designed and developed in his leisure time the 
	 functions and the complex dialog box of the RSA cryptosystem, 
	 which enable you to execute the RSA samples of this article.}

>>>>

The  computations for the factorization of RSA-160 took place at the 
German Information Security Agency (GISA) in Bonn\footnote{%
Every year the GISA\index{GISA} creates a paper to describe which 
crypto algorithms are feasible to generate digital signatures according
to the German signature law -- under participation of experts from
economy and science. To review signature methods based on the
factorisation problem the GISA also co-operates with researchers from the
University of Bonn.
Further information about crypto algorithms can be found on the web page of GISA:
   \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
{\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}}.
}.

@@

  \item {Mr.\ Henrik Koy for making many very useful suggestions, 
         for the very constructive proof-reading this article and
         for helping with TeX. \\
	 Mr.\ Koy designed and developed in his leisure time the 
	 functions and the complex dialog box of the RSA cryptosystem, 
	 which enables you to execute the RSA samples of this article.}

### 2003-07-07 script/en/chapters/primes.tex
*** empty log message ***

Almost all known huge prime numbers are special candidates, called
\index{Mersenne Marin} {\em Mersenne numbers}\footnote{%
Marin Mersenne, French priest and mathematician, Sep 08, 1588 -- Sep 01, 1648.
\index{Mersenne Marin}
}
of the form $2^p -1,$ where $p$ is
a prime. Not all Mersenne numbers are prime:

@@

Mersenne claimed that $2^{67}-1$ is a prime number. There is also a mathematical
history behind this claim: it first took over 200 years before 
\index{Lucas Edouard} Edouard Lucas (1842-1891) proved that this number 
is composite.
However, he argued indirectly and did not name any of the factors. Then Frank
Nelson showed in 1903 which factors make up this composite number: 
$$ 2^{67} -1
=147, 573, 952, 589, 676, 412, 927 = 193, 707, 721 \cdot 761, 838, 257, 287. $$
He admitted to having worked 20 years on the factorisation 
\index{Factorisation} (expression as a product of prime factors)\footnote{%
Using CrypTool\index{CrypTool} v1.3 you can factorize numbers in the 
following way: menu {\bf Indiv. Procedures \textbackslash{} RSA Demonstration 
\textbackslash{} Factorisation of a Number}.
}
of this 21-digit decimal number!

@@

As early as in the 19th century, it was discovered that the $29$-digit number $$
f(7) = 2^{2^7} + 1 $$ is not prime. However, it was not until 1970 that
Morrison/Billhart managed to factorise it.
\begin{eqnarray*}
f(7) & = & 340,282,366,920,938,463,463,374,607,431,768,211,457 \\
& = & 59, 649, 589, 127, 497, 217 \cdot  5,704,689,200,685,129,054,721
\end{eqnarray*}\label{f7}

@@

   This is a another possible generalisation of the Mersenne 
   numbers\index{Mersenne!number!generalized}.
   The \index{Cunningham project} \textbf{Cunningham project} determines
   the factors of all composite numbers that are formed as follows:
   $$ f(b,n) = b^n \pm 1  \quad {\rm for~} b = 2, 3, 5, 6, 7, 10, 11, 12 $$
   ($b$ is not equal to multiples of bases already used, such as $4, 8, 9$).

@@

\pagebreak %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --------------------------------------------------------------------------
\subsubsection{$f(n) = n^2 + n + 41$}
\label{L-Polynomfunktion01-41}
   This sequence starts off very {\em promisingly},  
   but is far from being a proof.

@@

% --------------------------------------------------------------------------
\subsubsection[Catalan's conjecture]{Catalan's conjecture\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, Belgian mathematician, May 5, 1814$-$Feb 14, 1894.\\
After him the so-called {\em Catalan numbers}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2 $ \\
$=  1, 2, 5, 14, 42, 132, 429, 1430, 
4862, 16796, 58786, 208012, 742900, 2674440, 9694845, ... $
are named.
    }
Catalan \index{Catalan Eugene}  \index{Number!Catalan}
conjectured that $ C_4 $ is a prime:
$$
 \begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
    (see {\href{http://www.utm.edu/research/primes/mersenne.shtml}{\tt
http://www.utm.edu/research/primes/mersenne.shtml}} under Conjectures and
Unsolved Problems).

>>>>

Almost all known huge prime numbers are special candidates, called
\index{Mersenne Marin} {\em Mersenne numbers}\footnote{%
Marin Mersenne, French priest and mathematician, Sep 08, 1588 $-$ Sep 01, 1648.
\index{Mersenne Marin}
}
of the form $2^p -1,$ where $p$ is
a prime. Not all Mersenne numbers are prime:

@@

Mersenne claimed that $2^{67}-1$ is a prime number. There is also a mathematical
history behind this claim: it first took over 200 years before 
\index{Lucas Edouard} Edouard Lucas (1842-1891) proved that this number 
is composite.
However, he argued indirectly and did not name any of the factors. Then Frank
Nelson Cole\index{Cole Frank Nelson}\footnote{%
Frank Nelson Cole, American mathematician, Sep. 20, 1861 $-$ May 26, 1926.}
showed in 1903 which factors make up this composite number: 
$$ 2^{67} -1
=147, 573, 952, 589, 676, 412, 927 = 193, 707, 721 \cdot 761, 838, 257, 287. $$
He admitted to having worked 20 years on the factorisation 
\index{Factorisation} (expression as a product of prime factors)\footnote{%
Using CrypTool\index{CrypTool} v1.3 you can factorize numbers in the 
following way: menu {\bf Indiv. Procedures \textbackslash{} RSA Demonstration 
\textbackslash{} Factorisation of a Number}.
}
of this 21-digit decimal number!

@@

As early as in the 19th century, it was discovered that the $29$-digit number $$
f(7) = 2^{2^7} + 1 $$ is not prime. However, it was not until 1970 that
Morrison/Billhart managed to factorise it.
\begin{eqnarray*}\label{f7}
f(7) & = & 340,282,366,920,938,463,463,374,607,431,768,211,457 \\
& = & 59, 649, 589, 127, 497, 217 \cdot  5,704,689,200,685,129,054,721
\end{eqnarray*}

@@

   This is another possible generalisation of the Mersenne 
   numbers\index{Mersenne!number!generalized}.
   The \index{Cunningham project} \textbf{Cunningham project} determines
   the factors of all composite numbers that are formed as follows:
   $$ f(b,n) = b^n \pm 1  \quad {\rm for~} b = 2, 3, 5, 6, 7, 10, 11, 12 $$
   ($b$ is not equal to multiples of bases already used, such as $4, 8, 9$).

@@

% --------------------------------------------------------------------------
\subsubsection{$f(n) = n^2 + n + 41$}
\label{L-Polynomfunktion01-41}
   This sequence starts off very {\em promisingly},  
   but is far from being a proof.

@@

% --------------------------------------------------------------------------
\subsubsection[Catalan's conjecture]{Catalan's conjecture\footnotemark}
    \footnotetext{%
Eugene Charles Catalan, Belgian mathematician, May 5, 1814$-$Feb 14, 1894.\\
After him the so-called {\em Catalan numbers}
$A(n) = (1 / (n+1) ) * (2n)! / (n!)^2 $ \\
$=  1, 2, 5, 14, 42, 132, 429, 1430, 
4862, 16796, 58786, 208012, 742900, 2674440, 9694845, ... $
are named.
    }
Catalan \index{Catalan Eugene}  \index{Number!Catalan}
conjectured that $ C_4 \;$ is a prime:
$$
 \begin{array}{l}
C_0 = 2, \\
C_1 = 2^{C_0} - 1,  \\
C_2 = 2^{C_1} - 1,  \\
C_3 = 2^{C_2} - 1, \\
C_4 = 2^{C_3} - 1, \cdots \\
\end{array}
$$
    (see {\href{http://www.utm.edu/research/primes/mersenne.shtml}{\tt
http://www.utm.edu/research/primes/mersenne.shtml}} under Conjectures and
Unsolved Problems).
